---
title       : Cornwall wind energy estimator
subtitle    : 
author      : Michael Hunt
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

Cornwall is England's windiest county.  
  
Many wind turbines have been sited here, and many more are planned.  

  
However the average wind speed varies with location and height above ground.


It is useful to project developers to have an idea of how much energy any model of wind turbine would produce annually wherever in the county it is placed, on however high a tower.
  

We have produced a Shiny app that uses wind speed atlas data for the county together with wind turbine specifications to produce a clickable map, where the annual energy production (AEP) of a turbine is caluclated, at whatever location in the county is selected, whatever the heighht of the turbine tower, for a range of turbines commmonly seen in the county.



--- 

```{r get the data,echo=FALSE}
# read in wind speed map of Cornwall taken from the UK NOABL wind speed atlas.
# wind speeds are given for 10 m, 25m and 45 m heights with 1 km2 resolution
windSpeeds<-read.csv("windspeeds.csv",stringsAsFactors=FALSE)


# get Cornwall boundary data
adm <- getData('GADM', country='GBR', level=2)
cornwall=adm[adm$NAME_2=="Cornwall",]
```
Here is the map of Cornwall

```{r map of Cornwall,echo=FALSE}
# get Cornwall boundary data
library(raster)
adm <- getData('GADM', country='GBR', level=2)
cornwall=adm[adm$NAME_2=="Cornwall",]
```

```{r plot map of Cornwall,cache=FALSE}
library(leaflet)
m <- leaflet() %>%
  addTiles() %>%  
  setView(lng = -4.7, lat = 50.4, zoom = 10)
m  
```




